---
title: "DESeq2_example"
author: "Jin"
date: '2022-06-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# BiocManager::install("DESeq2")
# BiocManager::install("pasilla")
```

## Standard Workflow

This is standard workflow following **DESeq2** tutorial.  
Vignettes see: <https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html>.

As input, DESeq2 expects un-normalized count data as obtained.  


```{r, echo = FALSE}
library(DESeq2)
```

## Construct a count matrix from single-cell data

The object class used by DESeq2 package to store the read counts and the intermediate estimated quantities during statistial analysis is the DESeqDataSet.  
A DESeqDataSet object must have an associated design formula. The design formula expresses the variables which will be used in modeling.  
DESeqDataSet can be constructed from a count matrix.  

```{r}
library(pasilla)
pasCts <- system.file("extdata",
                      "pasilla_gene_counts.tsv",
                      package="pasilla", mustWork=TRUE)
pasAnno <- system.file("extdata",
                       "pasilla_sample_annotation.csv",
                       package="pasilla", mustWork=TRUE)
cts <- as.matrix(read.csv(pasCts,sep="\t",row.names="gene_id"))
coldata <- read.csv(pasAnno, row.names=1)
coldata <- coldata[,c("condition","type")] # select metadata needed
coldata$condition <- factor(coldata$condition) # make factor
coldata$type <- factor(coldata$type) # make factor
```

Have a look at the cts and coldata:
```{r}
head(cts)
head(coldata)
```

Columns in the count matrix and rows of the column data should in the same order. So we need to rearrange this.
```{r}
rownames(coldata) <- sub("fb", "", rownames(coldata))
all(rownames(coldata) %in% colnames(cts))

all(rownames(coldata) == colnames(cts))

cts <- cts[, rownames(coldata)]
all(rownames(coldata) == colnames(cts))
```


